\texttt{clarify}: Simulation-Based Inference for Regression Models

### 2024-05-30 clarify/Submission 2/clarify.tex
Reorganized for resubmission

% !TeX root = RJwrapper.tex
\title{clarify: Simulation-Based Inference for Regression Models}

@@

\abstract{%
Simulation-based inference is an alternative to the delta method for computing the uncertainty around regression post-estimation (i.e., derived) quantities such as average marginal effects and adjusted predictions. It avoids several assumptions required by the delta method that are commonly violated, especially in nonlinear models, but does not require models to be refit as when using the bootstrap. It works by drawing model parameters from their joint distribution and estimating quantities of interest from each set of simulated values, which form the sampling distribution of the quantity. \pkg{clarify} provides a simple, unified interface for performing simulation-based inference for any user-specified derived quantities as well as wrappers for common quantities of interest. \pkg{clarify} supports a large and growing number of models through its interface with the \pkg{marginaleffects} package and provides native support for multiply imputed data.
}

@@

Although regression models are frequently used in empirical research to study relationships among variables, often the quantity of substantive interest is not one of the coefficients of the model, but rather a quantity derived from the coefficients, such as predicted values or average marginal effects. The usual method for estimating the uncertainty of the derived quantities is known as the ``delta method'', which involves two approximations: 1) that the variance of the derived quantity can be represented as a first-order Taylor series, and 2) that the resulting estimate is normally distributed\footnote{In addition, most software that uses the delta method uses numerical differentiation, which also involves an approximation.}. In many cases, especially with nonlinear models, these approximation can fail badly. \pkg{clarify} implements an alternative to the delta method---simulation-based inference---which involves simulating the sampling distributions of the derived quantities.

The methodology \pkg{clarify} relies on is described in King, Tomz, and Wittenberg (2000) and Rainey (2023). Simulation-based inference involves taking draws from a specified joint distribution of model parameters, computing derived quantities from these draws, and collecting the derived quantities in a ``posterior'' distribution, which is taken to be the sampling distribution of the derived quantity, and from which uncertainty measures (standard errors and confidence intervals) can be computed. This method assumes the model parameters are drawn from a multivariate Normal (or T) distribution with means at the estimated values and covariance equal to the asymptotic covariance matrix of the estimated values. Arriving at the posterior distribution does not require taking any derivatives or making any approximations except those used to define the sampling distribution of the parameter estimates. The delta method makes these same approximation in addition to the approximations of the variance and the shape of the sampling distribution of the estimates.

More formally, we fit a regression model \(y_i = f(x_i; \beta)\), such a linear or generalized linear model with model coefficients \(\beta\). We assume\[
\hat{\beta} \sim \text{MVN}(\beta, \Sigma_{\hat{\beta}})
\]where \(\hat{\beta}\) is the vector of estimates of \(\beta\) and \(\Sigma_{\hat{\beta}}\) is their asymptotic covariance matrix. We define a function \(\tau(\beta)\) that represents a quantity of interest derived from the model parameters, and compute its estimate \(\hat{\tau}(\beta)\) as \(\tau(\hat{\beta})\). Next, we take \(M\) draws \(\tilde{\beta}^{(j)}\) for \(j\in(1, \dots, M)\) from a multivariate Normal distribution with mean vector \(\boldsymbol{\mu} = \hat{\beta}\) and covariance \(\Sigma = \hat{\Sigma}_{\hat{\beta}}\), where \(\hat{\Sigma}_{\hat{\beta}}\) is an estimate of the asymptotic covariance matrix of the parameter estimates. We use the distribution of \(\tilde{\tau}=\tau(\tilde{\theta})\) as the ``posterior'' sampling distribution of \(\hat{\tau}(\beta)\), and compute its variance as\[
\sigma^2_{\hat{\tau}(\theta)} = \frac{1}{M}\sum^M_{j=1}{(\tilde{\tau}^{(j)} - \bar{\tilde{\tau}})^2}
\]and quantile \(100(1-\alpha)\%\) confidence interval limits as \(\left[\tilde{\tau}_{(\frac{\alpha}{2})}, \tilde{\tau}_{(1-\frac{\alpha}{2})}\right]\) where \(\tilde{\tau}_{(k)}\) is the \(k\)th value of \(\tilde{\tau}\) when arranged in ascending order.

Similar functionality exists in the \pkg{CLARIFY} package in Stata {[}Tomz, Wittenberg, and King (2003){]}\footnote{Despite the similar name, the R package \pkg{clarify} and the Stata package \pkg{CLARIFY} differ in several ways, one of which is that the estimates reported by \pkg{clarify} in R are those computed using the original model coefficients, whereas those reported by \pkg{CLARIFY} in Stata are those computed as the average of the simulated distribution. The R implementation avoids the ``simulation-induced bias'' described by Rainey (2023).} and used to be available in the \texttt{Zelig} R package (Imai, King, and Lau 2008), though there are differences in these implementations. \pkg{clarify} provides additional flexibility by allowing the user to request any derived quantity, in addition to providing shortcuts for common quantities, including predictions at representative values, average marginal effects, and average dose-response functions (described below). \pkg{clarify} relies on and can be seen as a companion to the \CRANpkg{marginaleffects} package, which offers similar functionality but primarily uses the delta method for calculating uncertainty (though simulation-based inference is supported in a more limited capacity as well).

\hypertarget{using}{%
\subsection{\texorpdfstring{Using \pkg{clarify}}{Using }}\label{using}}

There are four steps to using \pkg{clarify}:

@@

In the sections below, we'll describe how to implement these steps in detail. First, we'll load \pkg{clarify} using \texttt{library()}.

@@

For a running example, we'll use the \texttt{lalonde} dataset in the \pkg{MatchIt} package (Ho et al. 2011), which contains data on 614 participants enrolled in a job training program or sampled from a survey (Dehejia and Wahba 1999). The treatment variable is \texttt{treat} and the outcome is \texttt{re78}, and all other variables are confounders. Although the original use of this dataset was to estimate the effect of \texttt{treat} on \texttt{re78}, we'll use it more generally to demonstrate all of \pkg{clarify}'s capabilities. In addition, we'll use a transformation of the outcome variable to demonstrate applications to nonlinear models, for which the benefits of simulation-based inference are more apparent.

@@

If one's model is not supported by \pkg{clarify}, one can omit the \texttt{fit} argument and just specify the \texttt{vcov} and \texttt{coefs} argument, which will draw the coefficients from the distribution named in \texttt{dist} (\texttt{"normal"} by default).

@@

For example, let's say our derived quantity of interest is the predicted probability of the outcome for participant PSID1. We would specified our \texttt{FUN} function as follows:

@@

The resulting \texttt{clarify\_est} object contains the simulated estimates in matrix form as well as the estimate computed on the original coefficients. We'll examine the sampling distribution shortly, but first we'll demonstrate computing a derived quantity from the coefficients directly.

@@

There are several functions in \pkg{clarify} that serve as convenience wrappers for \texttt{sim\_apply()} to automate some common derived quantities of interest. These include

@@

\begin{itemize}
\item
  \texttt{x} -- the \texttt{clarify\_est} object (the output of a call to \texttt{sim\_apply()}).
\item
  \texttt{parm} -- the names or indices of the quantities to be plotted if more than one was estimated in \texttt{sim\_apply()}; if unspecified, all will be plotted.
\item
  \texttt{ci} -- whether to display lines at the uncertainty bounds. The default is \texttt{TRUE} to display them.
\item
  \texttt{level} -- if \texttt{ci} is \texttt{TRUE}, the desired two-sided confidence level. The default is .95 so that that the bounds are at the .025 and .975 quantiles when \texttt{method} (see below) is \texttt{"quantile"}.
\item
  \texttt{method} -- if \texttt{ci} is \texttt{TRUE}, the method used to compute the bounds. Allowable methods include a Normal approximation (\texttt{"wald"}) or using the quantiles of the resulting distribution (\texttt{"quantile"}). The Normal approximation involves multiplying the standard deviation of the estimates (i.e., which functions like the standard error of the sampling distribution) by the critical Z-statistic computed using \texttt{(1-level)/2} to create a symmetric margin of error around the point estimate. The default is \texttt{"quantile"} to instead use quantile-based bounds, which are more appropriate when the distribution is non-Normal. However, quantile-based bounds may require more simulations to stabilize.
\item
  \texttt{reference} -- whether to display a normal density over the plot for each estimate. The default is \texttt{FALSE} to omit them.
\end{itemize}

@@

From the plot, one can see that the distribution of simulated values is non-Normal, asymmetrical, and not centered around the estimate, with no values falling above 1 because the outcome is a predicted probability. Given its non-Normality, the quantile-based bounds are clearly more appropriate than those resulting from the Normal approximation, as the bounds computed from the Normal approximation would be outside the bounds of the estimate. The plot itself is a \texttt{ggplot} object that can be modified using \texttt{ggplot2} syntax.

@@

We'll use \texttt{summary()} with the default arguments on our first \texttt{clarify\_est} object to view the point estimate and quantile-based uncertainty bounds.

@@

\hypertarget{sim_apply-wrappers-sim_setx-sim_ame-sim_adrf}{%
\subsection{\texorpdfstring{\texttt{sim\_apply()} wrappers: \texttt{sim\_setx()}, \texttt{sim\_ame()}, \texttt{sim\_adrf()}}{sim\_apply() wrappers: sim\_setx(), sim\_ame(), sim\_adrf()}}\label{sim_apply-wrappers-sim_setx-sim_ame-sim_adrf}}

\texttt{sim\_apply()} can be used to compute the simulated sampling distribution for an arbitrary derived quantity of interest, but there are some quantities that are common in applied research and may otherwise be somewhat challenging to program by hand, so \pkg{clarify} provides shortcut functions to make computing these quantities simple. These functions include \texttt{sim\_setx()}, \texttt{sim\_ame()}, and \texttt{sim\_adrf()}. Each of these can be used only when regression models compatible with \pkg{clarify} are supplied to the original call to \texttt{sim()}.

@@

\texttt{sim\_setx()} provides an interface to compute predictions at representative and user-supplied values of the predictors. For example, we might want to know what the effect of treatment is for a ``typical'' individual, which corresponds to the contrast between two model-based predictions (i.e., one under treatment and one under control for a unit with ``typical'' covariate values). This functionality mirrors the \texttt{setx()} and \texttt{setx1()} functionality of \texttt{Zelig} (which is where its name originates) and provides similar functionality to functions in \pkg{modelbased}, \pkg{emmeans}, \pkg{effects}, and \pkg{ggeffects}.

For each predictor, the user can specify whether they want predictions at specific values or at ``typical'' values, which are defined in \pkg{clarify} as the mode for unordered categorical and binary variables, the median for ordered categorical variables, and the mean for continuous variables. Predictions for multiple predictor combinations can be requested by specifying values that will be used to create a grid of predictor values, or the grid itself can be supplied as a data frame of desired predictor profiles. In addition, the ``first difference'', defined here as the difference between predictions for two predictor combinations, can be computed.

@@

Here, we'll use \texttt{sim\_setx()} to examine predicted values of the outcome for control and treated units, at \texttt{re75} set to 0 and 20000, and \texttt{race} set to ``black''.

@@

We can plot the distributions of the simulated values using \texttt{plot()}, which also separates the predictions by the predictor values (it's often clearer without the uncertainty bounds). The \texttt{var} argument controls which variable is used for faceting the plots.

@@

We'll return to display average dose-response functions using \texttt{sim\_adrf()} later.

@@

Using predicted values and effects at representative values is one way to summarize regression models, but another way is to compute average adjusted predictions (AAPs), contrasts of AAPs, and average marginal effects (AMEs). The definitions for these terms may vary and the names for these concepts differ across sources, but here we define AAPs as the average of the predicted values for all units after setting one predictor to a chosen value, and we define AMEs for binary predictors as the contrast of two AAPs and for continuous predictors as the average of instantaneous rate of change in the AAP corresponding to a small change in the predictor from its observed values across all units\footnote{In \pkg{marginaleffects}, AAPs are computed using \texttt{avg\_predictions()}, AMEs for binary variables are computed using \texttt{avg\_comparisons()}, and AMEs for continuous variables are computed using \texttt{avg\_slopes()}. AAPs are sometimes known as average ``counterfactual'' predictions.}.

@@

Here, we'll use \texttt{sim\_ame()} to compute the AME of \texttt{treat} just among those who were treated (in causal inference, this is known as the average treatment effect in the treated, or ATT (Greifer and Stuart 2023)). We'll request our estimate to be on the risk ratio scale.

@@

We can use \texttt{summary()} to display the estimates and their uncertainty intervals. Here, we'll also use \texttt{null} to include a test for the null hypothesis that the risk ratio is equal to 1.

@@

We can compute the AME for a continuous predictor. Here, we'll consider \texttt{age} (just for demonstration; this analysis does not have a valid interpretation).

@@

Below, we'll examine effect modification of the ATT by the predictor \texttt{married} using the \texttt{by} argument to estimate AAPs and their ratio within levels of \texttt{married}:

@@

Here, we'll consider \texttt{age} (just for demonstration; this analysis does not have a valid interpretation) and compute the ADRF and AMEF of \texttt{age} on the outcome. We'll only examine ages between 18 and 50, even though the range of \texttt{age} goes slightly beyond these values. First, we'll compute the ADRF of \texttt{age}, which examines how the outcome would vary on average if one set all unit's value of \texttt{age} to each value between 18 and 50 (here we only use even ages to speed up computation).

@@

Next we'll compute the AMEF, the effect of \texttt{age} at each level of \texttt{age}.

@@

\hypertarget{transforming-and-combining-estimates}{%
\subsection{Transforming and combining estimates}\label{transforming-and-combining-estimates}}

@@

Similarly, to compute a derived or transformed quantity from a \texttt{clarify\_est} object, we can use \texttt{transform()}. Here, we'll compute the risk difference AME of \texttt{treat}; previously, we used \texttt{sim\_ame()} to compute the AAPs and the risk ratio.

@@

When we run \texttt{summary()} on the output, the new quantity, which we named ``RD'', will be displayed along with the other estimates. We'll also set a null value for this quantity.

@@

\texttt{cbind()} is another generic R function that is typically used to combine two or more datasets columnwise (i.e., to widen a dataset). In \pkg{clarify}, \texttt{cbind()} can be used to combine two \texttt{clarify\_est} objects so that the estimates can be examined jointly and so that it is possible to compare them directly. For example, let's say we computed AMEs in two subgroups using \texttt{subset} and wanted to compare them. To do so, we call \texttt{sim\_ame()} twice, one for each subset (though in practice it is more effective to use \texttt{by}; this is just for illustration):

@@

Here, we computed the risk difference for the subgroups \texttt{race\ =\ "black"} and \texttt{race\ =\ "hispan"}. If we wanted to compare the risk differences, we could combine them and compute a new quantity equal to their difference. We'll do that below.

First, we need to rename to quantities in each object so they don't overlap; we can do so using \texttt{names()}, which has a special method for \texttt{clarify\_est} objects.

@@

\hypertarget{using-with-multiply-imputed-data}{%
\subsection{\texorpdfstring{Using \pkg{clarify} with multiply imputed data}{Using  with multiply imputed data}}\label{using-with-multiply-imputed-data}}

@@

Using \pkg{clarify} with multiply imputed data is simple. Rather than using \texttt{sim()}, we use the function \texttt{misim()}. \texttt{misim()} functions just like \texttt{sim()} except that it takes in a list of model fits (i.e., containing a model fit to each imputed dataset) or an object containing such a list (e.g., a \texttt{mira} object from \texttt{mice::with()} or a \texttt{mimira} object from \texttt{MatchThem::with()}). \texttt{misim()} simulates coefficient distributions within each imputed dataset and then appends them together to a form a single combined set of coefficient draws.

\texttt{sim\_apply()} and its wrappers accept the output of \texttt{misim()} and compute the desired quantity using each set of coefficients. When these function rely on using a dataset (e.g., \texttt{sim\_ame()}, which averages predicted outcomes across all units in the dataset used to fit the model), they automatically know to associate a given coefficient draw with the imputed dataset that was used to fit the model that produced that draw. In user-written functions supplied to the \texttt{FUN} argument of \texttt{sim\_apply()}, it is important to correctly extract the dataset from the model fit. This is demonstrated below.

@@

Below we illustrate using \texttt{misim()} and \texttt{sim\_apply()} with multiply imputed data. We'll use the \texttt{africa} dataset from the \pkg{Amelia} package.

@@

The function we'll be applying to each imputed dataset will be one that computes the average marginal effect of \texttt{infl}. (We'll run the same analysis afterward using \texttt{sim\_ame()}.)

@@

Note that \texttt{misim()} is compatible with model fit objects from \pkg{mice}, \pkg{Amelia}, \pkg{MatchThem}, and any other package that produces a list of model fit objects with each corresponding to the output of a model fit to an imputed dataset.

\hypertarget{comparison-to-other-packages}{%
\subsection{Comparison to other packages}\label{comparison-to-other-packages}}

Several packages offer methods for computing interpretable quantities form regression models, including \pkg{emmeans}, \pkg{margins}, \pkg{modelbased}, and \pkg{marginaleffects}. Many of the quantities computed by these packages can also be computed by \pkg{clarify}, the primary difference being that \pkg{clarify} uses simulation-based inference rather than delta method-based inference.

>>>>

% !TeX root = RJwrapper.tex
\title{\texttt{clarify}: Simulation-Based Inference for Regression Models}

@@

\abstract{%
Simulation-based inference is an alternative to the delta method for computing the uncertainty around regression post-estimation (i.e., derived) quantities such as average marginal effects and adjusted predictions. It avoids several assumptions required by the delta method that are commonly violated, especially in nonlinear models, but does not require models to be refit as when using the bootstrap. It works by drawing model parameters from their joint distribution and estimating quantities of interest from each set of simulated values, which form the sampling distribution of the quantity. \texttt{clarify} provides a simple, unified interface for performing simulation-based inference for any user-specified derived quantities as well as wrappers for common quantities of interest. \texttt{clarify} supports a large and growing number of models through its interface with the \texttt{marginaleffects} package and provides native support for multiply imputed data.
}

@@

Although regression models are frequently used in empirical research to study relationships among variables, often the quantity of substantive interest is not one of the coefficients of the model, but rather a quantity derived from the coefficients, such as predicted values or average marginal effects. The usual method for estimating the uncertainty of the derived quantities is known as the ``delta method'', which involves two approximations: 1) that the variance of the derived quantity can be represented as a first-order Taylor series, and 2) that the estimate of the derived quantity is normally distributed\footnote{In addition, most software that uses the delta method uses numerical differentiation, which also involves an approximation.}. In many cases, especially with nonlinear models, these approximation can fail badly. For example, predicted probabilities close to 0 or 1 or ratios of coefficients or probabilities typically do not have normal (or even symmetrical) distributions in finite samples, and the usual Wald-type confidence interval limits produced from delta method standard errors can include values outside the domain of the quantity of interest.

\CRANpkg{clarify} implements an alternative to the delta method---simulation-based inference---which involves simulating the sampling distributions of the derived quantities. Simulation-based inference is not only often more accurate than using the delta method, it is also simpler to understand and implement, as it does not require understanding Taylor series or the calculus that underlies it. This makes it more palatable to nontechnical audiences and easier to learn for students without sacrificing statistical performance.

The methodology \CRANpkg{clarify} relies on is described in King, Tomz, and Wittenberg (2000) and Rainey (2023). Simulation-based inference involves taking draws from a specified joint distribution of model parameters, computing derived quantities from these draws, and collecting the derived quantities in a ``posterior'' distribution, which is taken to be the sampling distribution of the derived quantity, and from which uncertainty measures (standard errors and confidence intervals) can be computed. This method assumes the model parameters are drawn from a multivariate Normal (or T) distribution with means at the estimated values and covariance equal to the asymptotic covariance matrix of the estimated values, a standard assumption motivated by the central limit theorem that underlies usual inference on the original model parameters. Arriving at the posterior distribution does not require taking any derivatives or making any approximations beyond those usually used for inference on model parameter estimates. The delta method makes these same assumptions in addition to the approximations about the variance and the shape of the sampling distribution of the estimates of the derived quantities.

The nonparametric bootstrap is another alternative to the delta method for inference that does not require these approximations; it typically involves re-sampling individuals from the sample, fitting the model in each sample, and computing the quantity of interest from each model. Although bootstrapping tends to work well in practice, especially for complex and non-Normal estimators, refitting the model repeatedly can be prohibitively time-consuming and computationally expensive, especially for complicated models or large datasets. Simulation-based inference only requires the model to be fit once, and the simulations involve taking draws from a distribution produced from the single set of estimated parameters, making it much quicker in practice and allowing the user to capitalize on the already valid estimation of the model parameters.

More formally, we fit a regression model \(y_i = f(x_i; \beta)\), such as a linear or other generalized linear model with model coefficients \(\beta\). We assume\[
\hat{\beta} \sim \text{MVN}(\beta, \Sigma_{\hat{\beta}})
\]where \(\hat{\beta}\) is the vector of estimates of \(\beta\) and \(\Sigma_{\hat{\beta}}\) is their asymptotic covariance matrix. We define a function \(\tau(\beta)\) that represents a quantity of interest derived from the model parameters, and compute its estimate \(\widehat{\tau(\beta)}\) as \(\tau(\hat{\beta})\). Next, we take \(M\) draws \(\tilde{\beta}^{(j)}\) for \(j\in(1, \dots, M)\) from a multivariate Normal distribution with mean vector \(\mu = \hat{\beta}\) and covariance \(\Sigma = \hat{\Sigma}_{\hat{\beta}}\), where \(\hat{\Sigma}_{\hat{\beta}}\) is an estimate of the asymptotic covariance matrix of the parameter estimates. We use the distribution of \(\tilde{\tau}=\tau(\tilde{\theta})\) as the ``posterior'' sampling distribution of \(\widehat{\tau(\beta)}\), and compute its variance as\[
\sigma^2_{\widehat{\tau(\theta)}} = \frac{1}{M}\sum^M_{j=1}{(\tilde{\tau}^{(j)} - \bar{\tilde{\tau}})^2}
\]and quantile \(100(1-\alpha)\%\) confidence interval limits as \(\left[\tilde{\tau}_{(\frac{\alpha}{2})}, \tilde{\tau}_{(1-\frac{\alpha}{2})}\right]\) where \(\tilde{\tau}_{(k)}\) is the \(k\)th value of \(\tilde{\tau}\) when arranged in ascending order.

Similar functionality exists in the \pkg{CLARIFY} package in Stata\footnote{Despite the similar name, the R package \CRANpkg{clarify} and the Stata package \pkg{CLARIFY} differ in several ways, one of which is that the estimates reported by \CRANpkg{clarify} in R are those computed using the original model coefficients, whereas those reported by \pkg{CLARIFY} in Stata are those computed as the average of the simulated distribution. The R implementation avoids the ``simulation-induced bias'' described by Rainey (2023).} (Tomz, Wittenberg, and King 2003) and used to be available in the \texttt{Zelig} R package (Imai, King, and Lau 2008), though there are differences in these implementations. \CRANpkg{clarify} provides additional flexibility by allowing the user to request any derived quantity, in addition to providing shortcuts for common quantities, including predictions at representative values, average marginal effects, and average dose-response functions (described below). \CRANpkg{clarify} relies on and can be seen as a companion to the \CRANpkg{marginaleffects} package, which offers similar functionality but primarily uses the delta method for calculating uncertainty (though simulation-based inference is supported in a more limited capacity as well).

\hypertarget{using}{%
\subsection{\texorpdfstring{Using \CRANpkg{clarify}}{Using }}\label{using}}

There are four steps to using \CRANpkg{clarify}:

@@

In the sections below, we will describe how to implement these steps in detail. First, we will load \CRANpkg{clarify} using \texttt{library()}.

@@

For a running example, we will use the \texttt{lalonde} dataset in the \CRANpkg{MatchIt} package (Ho et al. 2011), which contains data on 614 participants enrolled in a job training program or sampled from a survey (Dehejia and Wahba 1999). The treatment variable is \texttt{treat} and the outcome is \texttt{re78}, and all other variables are confounders. Although the original use of this dataset was to estimate the effect of \texttt{treat} on \texttt{re78}, we will use it more generally to demonstrate all of \CRANpkg{clarify}'s capabilities. In addition, we will use a transformation of the outcome variable to demonstrate applications to nonlinear models, for which the benefits of simulation-based inference are more apparent.

@@

If one's model is not supported by \CRANpkg{clarify}, one can omit the \texttt{fit} argument and just specify the \texttt{vcov} and \texttt{coefs} argument, which will draw the coefficients from the distribution named in \texttt{dist} (\texttt{"normal"} by default).

@@

For example, we will let our derived quantity of interest be the predicted probability of the outcome for participant PSID1. We would specified our \texttt{FUN} function as follows:

@@

The resulting \texttt{clarify\_est} object contains the simulated estimates in matrix form as well as the estimate computed on the original coefficients. We will examine the sampling distribution shortly, but first we will demonstrate computing a derived quantity from the coefficients directly.

@@

There are several functions in \CRANpkg{clarify} that serve as convenience wrappers for \texttt{sim\_apply()} to automate some common derived quantities of interest. These include

@@

\begin{itemize}
\item
  \texttt{x} -- the \texttt{clarify\_est} object (the output of a call to \texttt{sim\_apply()}).
\item
  \texttt{parm} -- the names or indices of the quantities to be plotted if more than one was estimated in \texttt{sim\_apply()}; if unspecified, all will be plotted.
\item
  \texttt{ci} -- whether to display lines at the uncertainty bounds. The default is \texttt{TRUE} to display them.
\item
  \texttt{level} -- if \texttt{ci} is \texttt{TRUE}, the desired two-sided confidence level. The default is .95 so that that the bounds are at the .025 and .975 quantiles when \texttt{method} (see below) is \texttt{"quantile"}.
\item
  \texttt{method} -- if \texttt{ci} is \texttt{TRUE}, the method used to compute the bounds. Allowable methods include a Normal approximation (\texttt{"wald"}) or using the quantiles of the resulting distribution (\texttt{"quantile"}). The Normal approximation involves multiplying the standard deviation of the estimates (i.e., which functions like the standard error of the sampling distribution) by the critical Z-statistic computed using \texttt{(1-level)/2} to create a symmetric margin of error around the point estimate. The default is \texttt{"quantile"} to instead use quantile-based bounds, which are more appropriate when the distribution is non-Normal. However, quantile-based bounds may require more simulations to stabilize.
\item
  \texttt{reference} -- whether to display a normal density over the plot for each estimate. The default is \texttt{FALSE} to omit it.
\end{itemize}

@@

Overlaid on the plot in red is a Normal distribution with the same mean and standard deviation as the simulated values; this is requested by setting \texttt{reference\ =\ TRUE}. From the plot, one can see that the distribution of simulated values is non-Normal, asymmetrical, and not centered around the estimate, with no values falling above 1 because the outcome is a predicted probability. Given its non-Normality, the quantile-based bounds are clearly more appropriate than those resulting from the Normal approximation, as the bounds computed from the Normal approximation would be outside the bounds of the estimate. The plot itself is a \texttt{ggplot} object that can be modified using \texttt{ggplot2} syntax.

@@

We can use \texttt{summary()} with the default arguments on our first \texttt{clarify\_est} object to view the point estimate and quantile-based uncertainty bounds.

@@

\hypertarget{sim_apply-wrappers-sim_setx-sim_ame-sim_adrf}{%
\subsection{\texorpdfstring{\texttt{sim\_apply()} Wrappers: \texttt{sim\_setx()}, \texttt{sim\_ame()}, \texttt{sim\_adrf()}}{sim\_apply() Wrappers: sim\_setx(), sim\_ame(), sim\_adrf()}}\label{sim_apply-wrappers-sim_setx-sim_ame-sim_adrf}}

\texttt{sim\_apply()} can be used to compute the simulated sampling distribution for an arbitrary derived quantity of interest, but there are some quantities that are common in applied research and may otherwise be somewhat challenging to program by hand, so \CRANpkg{clarify} provides shortcut functions to make computing these quantities simple. These functions include \texttt{sim\_setx()}, \texttt{sim\_ame()}, and \texttt{sim\_adrf()}. Each of these can be used only when regression models compatible with \CRANpkg{clarify} are supplied to the original call to \texttt{sim()}.

@@

\texttt{sim\_setx()} provides an interface to compute predictions at representative and user-supplied values of the predictors. For example, we might want to know what the effect of treatment is for a ``typical'' individual, which corresponds to the contrast between two model-based predictions (i.e., one under treatment and one under control for a unit with ``typical'' covariate values). This functionality mirrors the \texttt{setx()} and \texttt{setx1()} functionality of \texttt{Zelig} (which is where its name originates) and provides similar functionality to functions in \CRANpkg{modelbased}, \CRANpkg{emmeans}, \CRANpkg{effects}, and \CRANpkg{ggeffects}.

For each predictor, the user can specify whether they want predictions at specific values or at ``typical'' values, which are defined in \CRANpkg{clarify} as the mode for unordered categorical and binary variables, the median for ordered categorical variables, and the mean for continuous variables. Predictions for multiple predictor combinations can be requested by specifying values that will be used to create a grid of predictor values, or the grid itself can be supplied as a data frame of desired predictor profiles. In addition, the ``first difference'', defined here as the difference between predictions for two predictor combinations, can be computed.

@@

Here, we will use \texttt{sim\_setx()} to examine predicted values of the outcome for control and treated units, at \texttt{re75} set to 0 and 20000, and \texttt{race} set to ``black''.

@@

We can plot the distributions of the simulated values using \texttt{plot()}, which also separates the predictions by the predictor values (it is often clearer without the uncertainty bounds). The \texttt{var} argument controls which variable is used for faceting the plots.

@@

We will return to display average dose-response functions using \texttt{sim\_adrf()} later.

@@

Using predicted values and effects at representative values is one way to summarize regression models, but another way is to compute average adjusted predictions (AAPs), contrasts of AAPs, and average marginal effects (AMEs). The definitions for these terms may vary and the names for these concepts differ across sources, but here we define AAPs as the average of the predicted values for all units after setting one predictor to a chosen value, and we define AMEs for binary predictors as the contrast of two AAPs and for continuous predictors as the average of instantaneous rate of change in the AAP corresponding to a small change in the predictor from its observed values across all units\footnote{In \CRANpkg{marginaleffects}, AAPs are computed using \texttt{avg\_predictions()}, AMEs for binary variables are computed using \texttt{avg\_comparisons()}, and AMEs for continuous variables are computed using \texttt{avg\_slopes()}. AAPs are sometimes known as average ``counterfactual'' predictions.} (Long and Freese 2014).

@@

Here, we will use \texttt{sim\_ame()} to compute the AME of \texttt{treat} just among those who were treated (in causal inference, this is known as the average treatment effect in the treated, or ATT (Greifer and Stuart 2023)). We will request our estimate to be on the risk ratio scale.

@@

We can use \texttt{summary()} to display the estimates and their uncertainty intervals. Here, we will also use \texttt{null} to include a test for the null hypothesis that the risk ratio is equal to 1.

@@

We can compute the AME for a continuous predictor. Here, we will consider \texttt{age} (just for demonstration; this analysis does not have a valid interpretation).

@@

Below, we will examine effect modification of the ATT by the predictor \texttt{married} using the \texttt{by} argument to estimate AAPs and their ratio within levels of \texttt{married}:

@@

Here, we will consider \texttt{age} (just for demonstration; this analysis does not have a valid interpretation) and compute the ADRF and AMEF of \texttt{age} on the outcome. We will only examine ages between 18 and 50, even though the range of \texttt{age} goes slightly beyond these values. First, we will compute the ADRF of \texttt{age}, which examines how the outcome would vary on average if one set all unit's value of \texttt{age} to each value between 18 and 50 (here we only use even ages to speed up computation).

@@

Next we will compute the AMEF, the effect of \texttt{age} at each level of \texttt{age}.

@@

\hypertarget{transforming-and-combining-estimates}{%
\subsection{Transforming and Combining Estimates}\label{transforming-and-combining-estimates}}

@@

Similarly, to compute a derived or transformed quantity from a \texttt{clarify\_est} object, we can use \texttt{transform()}. Here, we will compute the risk difference AME of \texttt{treat}; previously, we used \texttt{sim\_ame()} to compute the AAPs and the risk ratio.

@@

When we run \texttt{summary()} on the output, the new quantity, which we named ``RD'', will be displayed along with the other estimates. We will also set a null value for this quantity.

@@

\texttt{cbind()} is another generic R function that is typically used to combine two or more datasets columnwise (i.e., to widen a dataset). In \CRANpkg{clarify}, \texttt{cbind()} can be used to combine two \texttt{clarify\_est} objects so that the estimates can be examined jointly and so that it is possible to compare them directly. For example, if we were to compute AMEs in two subgroups using \texttt{subset} and wanted to compare them, we would call \texttt{sim\_ame()} twice, one for each subset (though in practice it is more effective to use \texttt{by}; this is just for illustration), as demonstrated below:

@@

Here, we computed the risk difference for the subgroups \texttt{race\ =\ "black"} and \texttt{race\ =\ "hispan"}. If we wanted to compare the risk differences, we could combine them and compute a new quantity equal to their difference. We will do that below.

First, we need to rename to quantities in each object so they do not overlap; we can do so using \texttt{names()}, which has a special method for \texttt{clarify\_est} objects.

@@

\hypertarget{using-with-multiply-imputed-data}{%
\subsection{\texorpdfstring{Using \CRANpkg{clarify} with Multiply Imputed Data}{Using  with Multiply Imputed Data}}\label{using-with-multiply-imputed-data}}

@@

Using \CRANpkg{clarify} with multiply imputed data is simple. Rather than using \texttt{sim()}, we use the function \texttt{misim()}. \texttt{misim()} functions just like \texttt{sim()} except that it takes in a list of model fits (i.e., containing a model fit to each imputed dataset) or an object containing such a list (e.g., a \texttt{mira} object from \texttt{mice::with()} or a \texttt{mimira} object from \texttt{MatchThem::with()}). \texttt{misim()} simulates coefficient distributions within each imputed dataset and then appends them together to a form a single combined set of coefficient draws.

\texttt{sim\_apply()} and its wrappers accept the output of \texttt{misim()} and compute the desired quantity using each set of coefficients. When these functions rely on using a dataset (e.g., \texttt{sim\_ame()}, which averages predicted outcomes across all units in the dataset used to fit the model), they automatically know to associate a given coefficient draw with the imputed dataset that was used to fit the model that produced that draw. In user-written functions supplied to the \texttt{FUN} argument of \texttt{sim\_apply()}, it is important to correctly extract the dataset from the model fit. This is demonstrated below.

@@

Below we illustrate using \texttt{misim()} and \texttt{sim\_apply()} with multiply imputed data. We will use the \texttt{africa} dataset from the \CRANpkg{Amelia} package.

@@

The function we will be applying to each imputed dataset will be one that computes the average marginal effect of \texttt{infl}. (We will run the same analysis afterward using \texttt{sim\_ame()}.)

@@

Note that \texttt{misim()} is compatible with model fit objects from \CRANpkg{mice}, \CRANpkg{Amelia}, \CRANpkg{MatchThem}, and any other package that produces a list of model fit objects with each corresponding to the output of a model fit to an imputed dataset.

\hypertarget{comparison-to-other-packages}{%
\subsection{Comparison to Other Packages}\label{comparison-to-other-packages}}

Several packages offer methods for computing interpretable quantities form regression models, including \CRANpkg{emmeans}, \CRANpkg{margins}, \CRANpkg{modelbased}, and \CRANpkg{marginaleffects}. Many of the quantities computed by these packages can also be computed by \CRANpkg{clarify}, the primary difference being that \CRANpkg{clarify} uses simulation-based inference rather than delta method-based inference.

\CRANpkg{marginaleffects} offers the most similar functionality to \CRANpkg{clarify}, and \CRANpkg{clarify} depends on functionality provided by \CRANpkg{marginaleffects} to accommodate a wide variety of regression models. \CRANpkg{marginaleffects} also offers simulation-based inference using \texttt{marginaleffects::inferences()} and support for arbitrary user-specified post-estimation functions using \texttt{marginaleffects::hypotheses()}. However, \CRANpkg{clarify} and \texttt{marignalefefcts} differ in several ways. The largest difference is that \CRANpkg{clarify} supports iterative building of more and more complex hypotheses through the \texttt{transform()} method, which quickly computes new quantities and transformation from the existing computed quantities, whereas \CRANpkg{marginaleffects} only supports a single transformation and, as of version 0.13.0, cannot use simulation-based inference for these quantities.
